---
title: "Forecasting Lending Amount in Australia Economy Through Bayesian VAR Approach"
author: "Zheyuan Li"

execute:
  echo: false
  
bibliography: references.bib
---
> This research project works for the assessment in ECOM90007 Macroeconometrics, which is 
the capstone subject in Master of Applied Econometrics in University of Melbourne.
>
> **Keywords.** bsvars, forecasting, loan market, lending rates, shrinkage, housing market

```{r,message=FALSE, warning=FALSE}
library(readabs)
library(readrba)
library(tseries)
library(tidyr)
library(dplyr)
library(forecast)
library(ggplot2)
library(lubridate)
library(httr)
library(readxl)
library(gridExtra)
response_1 <- GET("https://www.abs.gov.au/statistics/economy/price-indexes-and-inflation/residential-property-price-indexes-eight-capital-cities/dec-2021/641601.xlsx")
temp_file_1 <- tempfile(fileext = ".xlsx")
writeBin(content(response_1, "raw"), temp_file_1)
data <- read_excel(temp_file_1, sheet = 2)
data <- data[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data[[1]] <- origin_date + as.numeric(data[[1]]) - 2
dates = data[1]
rp_price_index <- data$`Residential Property Price Index ;  Weighted average of eight capital cities ;`
rp_price_index = as.numeric(rp_price_index)
rp_price_index <- log(rp_price_index)
plot_data <- data.frame(Date = dates, LogPriceIndex = rp_price_index)
```

```{r,message=FALSE, warning=FALSE}
response_2 <- GET("https://www.abs.gov.au/statistics/labour/employment-and-unemployment/labour-force-australia-detailed/feb-2024/6291014a.xlsx")
temp_file_2 <- tempfile(fileext = ".xlsx")
writeBin(content(response_2, "raw"), temp_file_2)
data_2 <- read_excel(temp_file_2, sheet = 2)
data_2 <- data_2[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_2[[1]] <- origin_date + as.numeric(data_2[[1]]) - 2
dates = data_2[1]
Unemployment <- data_2$`Unemployed total ;  Persons ;`
Unemployment = as.numeric(Unemployment)
Unemployment <- log(Unemployment)
plot_data_2 <- data.frame(Date = dates, LogPriceIndex = Unemployment)

```

```{r,message=FALSE, warning=FALSE}
response_3 <- GET("https://www.rba.gov.au/statistics/tables/xls/d03hist.xlsx?v=2024-04-02-00-51-45")
temp_file_3 <- tempfile(fileext = ".xlsx")
writeBin(content(response_3, "raw"), temp_file_3)
data_3 <- read_excel(temp_file_3, sheet = 1,skip = 1)
data_3 <- data_3[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_3[[1]] <- origin_date + as.numeric(data_3[[1]]) - 2
dates = data_3[1]
M1 <- data_3$M1
M1 = as.numeric(M1)
M1 <- log(M1)
plot_data_3 <- data.frame(Date = dates, M1_Amount = M1)
plot_data_3 <- na.omit(plot_data_3[c("Title", "M1_Amount")])
M1 = na.omit(M1)
```

```{r,message=FALSE, warning=FALSE}
response_4 <- GET("https://www.abs.gov.au/statistics/economy/price-indexes-and-inflation/producer-price-indexes-australia/dec-2023/642701.xlsx")
temp_file_4 <- tempfile(fileext = ".xlsx")
writeBin(content(response_4, "raw"), temp_file_4)
data_4 <- read_excel(temp_file_4, sheet = 2)
data_4 <- data_4[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_4[[1]] <- origin_date + as.numeric(data_4[[1]]) - 2
dates = data_4[1]
PPI <- data_4$`Index Numbers ;  Final ;  Total (Source) ;`
PPI = as.numeric(PPI)
PPI <- log(PPI)
plot_data_4 <- data.frame(Date = dates, PPI = PPI)
```

```{r,message=FALSE, warning=FALSE}
response_5 = GET("https://www.abs.gov.au/statistics/economy/price-indexes-and-inflation/consumer-price-index-australia/dec-quarter-2023/640101.xlsx")
temp_file_5 <- tempfile(fileext = ".xlsx")
writeBin(content(response_5, "raw"), temp_file_5)
data_5 <- read_excel(temp_file_5, sheet = 2)
data_5 <- data_5[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_5[[1]] <- origin_date + as.numeric(data_5[[1]]) - 2
dates = data_5[1]
CPI <- data_5$`Index Numbers ;  All groups CPI ;  Australia ;`
CPI = as.numeric(CPI)
CPI <- log(CPI)
plot_data_5 <- data.frame(Date = dates, CPI_Amount = CPI)
```

```{r,message=FALSE, warning=FALSE}
response_6 = GET("https://www.abs.gov.au/statistics/economy/national-accounts/australian-national-accounts-national-income-expenditure-and-product/dec-2023/5206002_Expenditure_Volume_Measures.xlsx")
temp_file_6 <- tempfile(fileext = ".xlsx")
writeBin(content(response_6, "raw"), temp_file_6)
data_6 <- read_excel(temp_file_6, sheet = 2)
data_6 <- data_6[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_6[[1]] <- origin_date + as.numeric(data_6[[1]]) - 2
dates = data_6[1]
GDP <- data_6$`GROSS DOMESTIC PRODUCT ;...42`
GDP = as.numeric(GDP)
GDP <- log(GDP)
plot_data_6 <- data.frame(Date = dates, GDP_Amount = GDP)
GDP = na.omit(GDP)
```

# Introduction
This project is focusing on the credit market in Australia. It intends to use Bayesian VAR approach in forecasting of lending amounts.

## The Research Objective and Question
The research aims to forecast lending indicators, specifically lending amounts, in Australia. Thus, the research question is: How can the lending indicators in Australia be accurately forecasted, and what are the resulting forecasts?

## Motivation of Research 
For banks and related financial institutions, they need to manage the risks associated with future interest rate fluctuations. Accurate lending amount forecasting can assist them in better assessing and managing risks, thereby safeguarding their assets and maintaining financial stability.

# Data and Properties
This section shows the data collection and their descriptive statistic including their series plotting and basic properties analyzing. 

## Variables Selection 
The selection of variables grounds in several macroeconomic indicators, derived from the official databases of the Australian Bureau of Statistics (ABS) and the Reserve Bank of Australia (RBA). The rationale for the choice of variables is informed by the work of scholars Marta Bańbura, Domenico Giannone, and Michele Lenza (2015) on European macroeconomic forecasting.

These indicators are closely linked to both loan rates and volumes. An increase in CPI and housing price indices signifies a decline in the purchasing power of money, necessitating a rise in loan rates to compensate for inflationary losses. Growth in GDP indicates economic expansion, leading to increased demand for capital, and consequently, higher loan volumes and rates. Higher unemployment rates suggest an economic downturn, prompting banks to lower interest rates to stimulate the economy. Additionally, an abundance of monetary supply leads banks to reduce loan rates to attract more borrowers.

Therefore, in my project, these following specific variables are used in BVAR model:


-   $M1_t$: M1 aggregate from RBA Database

-   $CPI_t$: CPI aggregate from ABS Database

-   $RPPI_t$: Residential Property Price Index aggregate from ABS Database

-   $GDP_t$: Expenditure on Gross Domestic Product (GDP) aggregate from ABS Database

-   $UNEMPLOY_t$: Unemplyemnt person aggregate from ABS Database 


```{r warning=FALSE}
p1 = ggplot(plot_data, aes(x = ...1, y = LogPriceIndex)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'RPPI')

p2 = ggplot(plot_data_2, aes(x = ...1, y = Unemployment)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'Unemploy')

p3 = ggplot(plot_data_3, aes(x = Title, y = M1_Amount)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'M1')

p4 = ggplot(plot_data_4, aes(x = ...1, y = PPI)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'PPI')

p5 = ggplot(plot_data_5, aes(x = ...1, y = CPI_Amount)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'CPI')

p6 = ggplot(plot_data_6, aes(x = ...1, y = GDP_Amount)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'GDP')

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2)
```


For the loan volumes to be forecasted, I have categorized them into housing loans and commercial loans, as they represent two distinct types of demand. The underlying logic influenced by macroeconomic factors might differ between them:

-   $House Lending_t$: House lending volume aggregate from RBA 

-   $Business Lending_t$: Business lending volume aggregate from RBA


```{r,message=FALSE, warning=FALSE}
response_7 = GET("https://www.rba.gov.au/statistics/tables/xls/d05hist.xlsx?v=2024-04-02-14-20-19")
temp_file_7 <- tempfile(fileext = ".xlsx")
writeBin(content(response_7, "raw"), temp_file_7)
data_7 <- read_excel(temp_file_7, sheet = 1, skip = 1)
data_7 <- data_7[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_7[[1]] <- origin_date + as.numeric(data_7[[1]]) - 2
dates = data_7[1]
Houselend <- data_7$`Lending to persons; Housing; Owner-occupiers`
Houselend = as.numeric(Houselend)
Houselend <- log(Houselend)
plot_data_7 <- data.frame(Date = dates, Houselend_Amount = Houselend)

```



```{r,message=FALSE, warning=FALSE}
response_8 = GET("https://www.rba.gov.au/statistics/tables/xls/d05hist.xlsx?v=2024-04-02-14-20-19")
temp_file_8 <- tempfile(fileext = ".xlsx")
writeBin(content(response_8, "raw"), temp_file_8)
data_8 <- read_excel(temp_file_8, sheet = 1, skip = 1)
data_8 <- data_8[-(1:9), ]
origin_date <- as.Date("1899-12-30")
data_8[[1]] <- origin_date + as.numeric(data_8[[1]]) - 2
dates = data_8[1]
Businesslend <- data_8$`Commercial lending; Business sector`
Businesslend = as.numeric(Businesslend)
Businesslend <- log(Businesslend)
plot_data_8 <- data.frame(Date = dates, Businesslend_Amount = Businesslend)
Businesslend = na.omit(Businesslend)
```



```{r, warning=FALSE}
p7 = ggplot(plot_data_7, aes(x = Title, y = Houselend_Amount)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'House Lend')


p8 = ggplot(plot_data_8, aes(x = Title, y = Businesslend_Amount)) +
  geom_line(color = 'blue') +
  theme_minimal() +
  labs(x = 'Date', y = 'Business Lend')

grid.arrange(p7,p8, ncol = 2)

```


```{r,warning=FALSE}
data_list <- list(
  rp_price_index = rp_price_index,
  Unemployment = Unemployment,
  M1 = M1,
  PPI = PPI,
  CPI = CPI,
  GDP = GDP,
  Houselend = Houselend,
  Businesslend = Businesslend
)

adf_results <- lapply(names(data_list), function(name) {
  adf_test_result <- adf.test(data_list[[name]], k = 5)
  data.frame(
    Test = name,
    Test_Statistic = adf_test_result$statistic,
    P_Value = adf_test_result$p.value,
    Lags = 5 
  )
})

adf_summary <- do.call(rbind, adf_results)
print(adf_summary)

```

```{r,warning=FALSE}
adf_diff_results <- lapply(names(data_list), function(name) {
  diff_series <- diff(data_list[[name]])
  adf_test_result <- adf.test(diff_series, k = 5)
  data.frame(
    Test = paste(name, "First Difference", sep = " - "),
    Test_Statistic = adf_test_result$statistic,
    P_Value = adf_test_result$p.value,
    Lags = 5
  )
})

adf_diff_summary <- do.call(rbind, adf_diff_results)

print(adf_diff_summary)
```


```{r}
par(mfrow=c(2, 4), mar=c(5, 2, 4, 2) + 0.1, mgp=c(3, 1, 0), cex.main=1, cex.axis=1, cex.lab=1)

for(name in names(data_list)) {
  ts_data <- data_list[[name]]
  Acf(ts_data, main = name)
}
```


```{r}
par(mfrow=c(2, 4), mar=c(5, 2, 4, 2) + 0.1, mgp=c(3, 1, 0), cex.main=1, cex.axis=1, cex.lab=1)

for(name in names(data_list)) {
  ts_data <- data_list[[name]]
  Pacf(ts_data, main = name)
}

```
## The Econometric Modelling

















